\include{header}
\begin{document}
\thispagestyle{empty}
{\bf Math 821, Spring 2014\\
Problem Set \#1\\
Due date: Friday, January 31}

As always, ``space'' means ``topological space'' and ``map'' means ``continuous function''.

\prob
Let $X$ be a path-connected space.
Prove that $X$ is connected.  (Recall that the
converse is not true --- the topologists' sine curve
is a counterexample.)

\prob Let $\Gamma$ be a finite graph.  (That is,
$\Gamma$, consists of finitely many copies of the closed interval
$I$, with some of the endpoints identified.
More precisely, $\Gamma=X/\sim$, where $X$ is the disjoint union of
finitely many copies $[a_1,b_1],\dots,[a_n,b_n]$ of
the closed unit interval $I$, and $\sim$ is an equivalence
relation in which every point $p\not\in\{a_1,b_1,\dots,a_n,b_n\}$
induces a singleton equivalence class.)  Prove that if $\Gamma$
is connected, then it is path-connected.  (In fact, this is true not merely
for graphs, but for all finite cell complexes.)

\prob Let $X$ and $Y$ be spaces and
let $f:X\surj Y$ be a map that is onto.
Prove that if $X$ is compact, then so is $Y$.

\prob [Hatcher p.18 \#1] Construct an explicit deformation retraction of the torus with one point deleted onto a graph consisting of two circles intersecting in a point.

\prob [Hatcher p.18 \#3, more or less]  (a) Show that the relation ``$X$ is homotopy equivalent to $Y$''
is an equivalence relation.

(b) Fix spaces $X,Y$ and let $f,g$ be maps $X\to Y$.  Show that the relation ``$f$ is homotopic to $g$'' is an equivalence relation.

(c) Show that a map homotopic to a homotopy equivalence is a homotopy equivalence.
More precisely:
Let $f:X\to Y$ be a homotopy equivalence.  Show that any map $g:X\to Y$ homotopic to $f$ is a homotopy equivalence.

\prob [Hatcher p.19 \#14] Given nonnegative integers $v,e,f$ with $v>0$, $f>0$, and $v-e+f=2$, construct a cell structure on $S^2$ having $v$ 0-cells, $e$ 1-cells, 
and $f$ 2-cells.  (Do not use any facts about spanning trees or Euler characteristic.)

\end{document}
