\include{header}
\begin{document}
\thispagestyle{empty}
{\bf Math 821, Spring 2014\\
Problem Set \#5\\
Due date: Friday, April 4}

\vfill\prob Verify that the simplicial boundary map defined by
\[\bd_n[v_0,\dots,v_n] = \sum_{i=0}^n (-1)^i [v_0,\dots,\widehat{v_i},\dots,v_n]\]
satisfies the equation $\bd_{n-1}\circ\bd_n=0$ for all $n$.  (Yes, this
calculation is done explicitly in Hatcher.
But it is so important that everyone should do it for themselves at least once.)

\vfill\prob Let $X$ be an abstract simplicial complex on vertex set $[n]$ and let $|X|$ be a geometric realization of $X$
(not necessarily the standard one --- it doesn't matter).
What invariant of $|X|$ corresponds to the dimension of $H^\Delta_0(X)$? 

\vfill\prob Consider the matrix
$$M=\begin{bmatrix} 1&1&0\\1&0&1\\0&1&1\end{bmatrix}.$$
Describe $\coker M$
(i) if $M$ is regarded as a linear transformation over $\Qq$;
(ii) if $M$ is regarded as a linear transformation over $\Zz$;
(iii) if $M$ is regarded as a linear transformation over $\Ff_q$
(the finite field with $q$ elements).

\vfill\prob [Hatcher p.131 \#4]  Compute \underline{by hand} the simplicial homology groups of the ``triangular parachute'' obtained from $\Delta^2$ by identifying its vertices to a single point.

\vfill\prob [Hatcher p.131 \#5]  Compute \underline{by hand} the simplicial homology groups of the Klein bottle using the $\Delta$-complex structure on p.102 (with two triangles).

\vfill\prob Check your answers on the last two problems using {\tt Macaulay2} or your favorite computer algebra system.
Submit your source code and output with your problem set. (For example, if you are using {\tt Macaulay2}, you can cut and paste the session transcript into your TeX file and use the {\tt verbatim} enviroment.)

Here is a link to  \href{http://www.jlmartin.faculty.ku.edu/math821/Macaulay2.pdf}{\fbox{how to get started with Macaulay2.}}

\vfill\prob Let $\Delta^{n,d}$ denote the $d$-skeleton of the $n$-simplex.  As an abstract simplicial complex, $\Delta$ is generated by all $(d+1)$-element subsets of $\{0,\dots,n\}$.  Use {\tt Macaulay2} (or another computer algebra system) to compute the homology groups of $\Delta^{n,d}$ for various values of $n$ and $d$.  Conjecture a general formula for $H_k(\Delta^{n,d})$ in terms of $n$, $d$ and $k$.  (Prove it, if you want.)
\end{document}
