\include{header}
\begin{document}
\thispagestyle{empty}
\begin{center}\large Math 409, Spring 2013\\
Final Exam Review Information
\end{center}

The final exam will be on \textbf{Monday, May 13, 1:30--4:00 PM, 301 Snow}
(the regular classroom).  Bring a pen or pencil and scratch paper.

The final exam will cover material from the entire course,
including the three main units: Euclidean geometry,
transformational geometry, and polyhedra.

\textbf{1. Euclidean geometry.}  Since the midterm was
all about Euclidean geometry, it won't be the main focus on the final
(roughly 25\% of the exam).
The problems on the final exam will be similar in difficulty
and subject to the ones on the midterm, and I'll give you the same list
of axioms and theorems.

\textbf{2. Transformational geometry.}  This was the big unit
of the second half of the course, so it will comprise roughly
50\% of the final exam.  Here is a list of
concepts you should understand and problems you should know how to solve.

\begin{itemize}

\item Understand the definition of transformation and group, and why the
following sets form groups:
(i) all transformations of space; (ii) all isometries of space;
(iii) all symmetries of an object in space (such as a polygon).
[Relevant problems: TG~5, TG~6]

\item Describe all symmetries of an object such as a polygon.
[Relevant problems: TG~1, TG~2, TG~13, TG~20]

\item Know the basic kinds of transformations (reflections, rotations,
translations) and how they interact.
[Relevant problems: TG~3, TG~4, TG~8, TG~9, TG~11, TG~12]

\item Be able to work with transformations in terms of their permutation words.
[Relevant problems: TG~16, TG~18, TG~20]

\item Construct and work with the multiplication table of a group.
[Relevant problems: TG~15, TG~16]

\item Know the Three-Point Theorem and the Three-Reflection Theorem.
[Relevant problems: TG~9, TG~10, TG~11]

\item Understand the group of symmetries of a regular polygon.
[Relevant problems: TG~18, TG~19]

\item Be able to count the symmetries of an object without having to list them all.
[Relevant problems: TG~17, TG~18, TG~19, TG~20, TG~21, TG~22]
\end{itemize}

\textbf{3. Polyhedra.}  This unit will comprise roughly 25\% of the
final exam.  Here's a list of concepts and problems.

\begin{itemize}
\item Understand how to use the ``handshaking formulas'' (see, e.g.,
Problem~9 in the notes on polyhedra).

\item Know Euler's formula and understand how to prove it (for instance,
by the ``fields, dikes and raging sea'' argument).

\item Know the classification of the Platonic solids,
and understand how to prove that they are the only ones.
\end{itemize}

Finally, the \href{http://www.jlmartin.faculty.ku.edu/math409/midterm-review.pdf}{study suggestions for the midterm exam} apply
equally well to the final exam.

\end{document}
