\include{header}
\begin{document}
\thispagestyle{empty}
{\bf Math 821 Problem Set \#1\\
Posted: Friday 1/28/11\\
Due date: Monday 2/7/11}

(Note: Some of these problems are proven as theorems in Munkres,
but try to do them yourself.)

\prob
Let $n$ and $m$ be nonnegative integers.  Let $X=\Rr^n$ and $X'=\Rr^m$, both equipped
with the standard topology.  Prove that the product topology on $X\x X'$
is exactly the standard topology on $\Rr^{n+m}$.

\prob
Let $(X,\TT)$ and $(X',\TT')$ be topological spaces,
let $\BB'$ be a basis for the topology $\TT'$,
and let $f:X\to X'$ be a function.  Prove that
$f$ is continuous if and only if $f^{-1}(U)$ is open 
for all $U\in\BB'$.

\prob
Let $X$ be a path-connected topological space.
Prove that $X$ is connected.  (Recall that the
converse is not true --- the topologists' sine curve
is a counterexample.)

\prob Let $\Gamma$ be a finite graph.  (That is,
$\Gamma$, consists of finitely many copies of the closed interval
$I$, with some of the endpoints identified.
More precisely, $\Gamma=X/\sim$, where $X$ is the disjoint union of
finitely many copies $[a_1,b_1],\dots,[a_n,b_n]$ of
the closed unit interval $I$, and $\sim$ is an equivalence
relation in which every point $p\not\in\{a_1,b_1,\dots,a_n,b_n\}$
induces a singleton equivalence class.)

Prove that if $\Gamma$
is connected, then it is path-connected.  (In fact, this is true not merely
for graphs, but for all finite cell complexes.)

\prob Let $X$ and $Y$ be topological spaces and
let $f:X\surj Y$ be a continuous function that is onto.
Prove that if $X$ is connected, then so is $Y$.

\prob Let $X$ and $Y$ be topological spaces and
let $f:X\surj Y$ be a continuous function that is onto.
Prove that if $X$ is compact, then so is $Y$.

\prob Let $n\geq 0$ be an integer, and let $X_n$ be the 
space you get by taking a strip of paper, twisting it $n$
times, and gluing the ends together.  (So $X_0$ is
a cylinder and $X_1$ is the M\"obius strip.)  For which
pairs $n,m$ are $X_n$ and $X_m$ homeomorphic?
\end{document}
