\include{header}
\usepackage{youngtab}
\begin{document}
\thispagestyle{empty}
{\bf Math 821 Problem Set \#2\\
Posted: Friday 2/11/11\\
Due date: Monday 2/21/11}

\prob (Hatcher, p.18, \#2)
Construct an explicit deformation retraction of $X=\Rr^n\sm\{0\}$
onto $S^{n-1}$.  (``Explicit'' means that you should write down an 
actual formula for the map $f_t:X\to X$, and check that the family of 
maps you have defined satisfies the conditions of a deformation 
retraction.)

\prob (Hatcher, p.19, \#12)
Show that a homotopy equivalence $f:X\to Y$ induces a bijection 
between the set of path-components of~$X$ and the set of path-components 
of~$Y$, and that~$f$ restricts to a homotopy equivalence from each 
path-component of~$X$ to the corresponding path-component of~$Y$. Prove 
also the corresponding statements with components instead of 
path-components. Deduce that if the components of a space~$X$ coincide 
with its path-components, then the same holds for any space~$Y$ homotopy 
equivalent to $X$.

\prob (Hatcher, p.19, \#17)
Construct a 2-dimensional cell complex that contains both an
annulus $S^1\x I$ and a M\"obius band as deformation retracts.
(This implies that these two spaces are homotopy-equivalent.)

\prob Let $X,Y$ be cell complexes of finite type.  Recall that the
\emph{$f$-polynomial} of~$X$ is
$$f(X;q) = \sum_{\text{cells } e_\alpha^i\in X} q^i
= \sum_{i=0}^{\dim X} f_i q^i$$
where $f_i$ is the number of $i$-dimensional cells (and recall that ``of finite type'' means that $f_i<\infty$ for each~$i$).
In terms of $f(X;q)$ and $f(Y;q)$,
find formulas for
\begin{enumerate}
\item[(a)] $f(X\x Y;q)$;
\item[(b)] $f(X/Y;q)$ (assuming that $(X,Y)$ is a CW-pair);
\item[(c)] $f(CX;q)$;
\item[(d)] $f(X*Y;q)$.
\end{enumerate}
You don't have to give detailed proofs.
(See pp.~8--10 of Hatcher for the definitions of these operations.)

\prob Let $0\leq k\leq n$.
Recall from class that the \emph{Grassmannian} $G(k,n)$
is defined as the space of $k$-dimensional subspaces $V\subset\Rr^n$,
so that in particular, $G(1,\Rr^n)=\Rr P^{n-1}$.  (Fact: Everything in
this problem works the same way if you change $\Rr$ to $\Cc$, except
that the dimensions of all the cells get doubled.)

\probpart Work out an explicit cell decomposition for $G(2,4)$
as a finite CW-complex.  That is,
describe how to decompose the set $G(2,4)$ into pieces, each of which is
isomorphic to a $\Rr$-vector space.
If you do this correctly (hint: row-reduced echelon form),
then the isomorphisms should be straightforward from the construction.

\probpart Describe the attaching poset of $G(2,4)$.  (Recall that this
is the partially ordered
set whose elements are the cells $e_\alpha$, and whose order
relation is given by $e_\alpha\geq e_\beta$ if
$\overline{e_\alpha}\supseteq e_\beta$).

\probpart A \emph{Ferrers diagram} is a collection of square boxes that
are top- and right-justified: for instance,
\begin{center}
\yng(6,6,4,3,1)
\end{center}
Write out the poset $P(2,2)$ of all Ferrers diagrams with at most two 
rows and at most two columns, ordered by containment (as sets of 
squares).  Compare it to your previous answer.

\probpart Make as many conjectures as you dare about the cell structure of
$G(k,n)$.  (In particular, what happens to the attaching poset if you
reverse all the relations?)

%\prob For $k\geq 1$, let $T_n$ denote the $n$-holed torus.
%
%\probpart Construct a cell complex structure on $T_n$.
%
%\probpart Show that $T_n$ is homotopy-equivalent to the cell complex
%$S^1\x(\bigvee_{i=1}^n S^1)$ (that is, take $n$ circles and wedge them together
%at a point to get a ``bouquet'', then take the product with another circle).

\prob (Extra credit; Hatcher, p.19, \#16)
Show that $S^\infty$ is contractible.

\end{document}
