\include{header}
\begin{document}
\bf Math 824, Fall 2012\\
Problem Set \#3\rm

{\bf Instructions:} Type up your solutions using LaTeX.  There is a
header file at\\
\href{http://www.jlmartin.faculty.ku.edu/math824/header.tex}{\tt http://www.math.ku.edu/$\sim$jmartin/math824/header.tex} with macros that may be useful.
E-mail me (\texttt{jmartin@math.ku.edu}) the PDF file under the name \textsl{$\{$your-name$\}$3.pdf}.

Deadline: {\bf 3:00 PM on Friday, October 5.}
\smallskip\hrule


\prob Determine, with proof, all pairs of integers $k\leq n$ such
that there exists a graph $G$ with $M(G)\isom U_k(n)$.  (Recall
that $U_k(n)$ is the matroid on $E=[n]$ such that every
subset of~$E$ of cardinality~$k$ is a basis.)

\prob Let $X$ and $Y$ be disjoint sets of vertices, and let $B$ be an $X,Y$-bipartite
graph: that is, every edge of $B$ has one endpoint in each of $X$ and $Y$.
For $V=\{x_1,\dots,x_n\}\subset X$, a \emph{transversal} of $V$ is a set $W=\{y_1,\dots,y_n\}\subset Y$
such that $x_iy_i$ is an edge of $B$.  (The set of all edges $x_iy_i$ is called a \emph{matching}.)
Let $\II$ be the family of all subsets of $X$
that have a transversal; it is immediate that $\II$ is a simplicial complex.

Prove that $\II$ is in fact a matroid independence system by verifying that the donation condition
holds.  (Suggestion: Write down an example
or two of a pair of independent sets $I,J$ with $|I|<|J|$, and use the corresponding matchings
to find a systematic way of choosing a vertex that $J$ can donate to $I$.)
These matroids are called \emph{transversal matroids}; along with linear and graphic matroids,
they are the other ``classical'' examples of matroids in combinatorics.)

\prob Let $G=(V,E)$ be a graph with $n$ vertices and $c$ components.
For a vertex coloring $f:V\to\Nn$, let $i(f)$ denote the number of ``improper'' edges,
i.e., whose endpoints are assigned the same color.
\emph{Crapo's coboundary polynomial} of $G$ is
  $$\bar\chi_G(q;t) ~=~ q^{-c}\sum_{f:V\to[q]} t^{i(f)}.$$
This is evidently a stronger invariant than the chromatic polynomial of $G$, which can be obtained as $q\bar\chi_G(q,0)$.
In fact, the coboundary polynomial provides the same information as the Tutte polynomial.

Prove that 
  $$\bar\chi_G(q;t) ~=~ (t-1)^{n-c} T_G\left(\frac{q+t-1}{t-1},t\right)$$
by finding a deletion/contraction recurrence for the coboundary polynomial.

\prob
Let $P$ be a chain-finite poset.
The \emph{kappa function} of $P$ is the element
of the incidence algebra $I(P)$ defined by $\kappa(x,y) = 1$
if $x\coveredby y$, $\kappa(x,y) = 0$ otherwise.

\probpart Give a condition on $\kappa$ that is equivalent   
to $P$ being ranked.

\probpart Give combinatorial interpretations of
$\kappa*\zeta$ and $\zeta*\kappa$.

(See next page for Problem \#5.)
\pagebreak

\prob
Let~$\Pi_n$ be the lattice of set partitions of~$[n]$.  Recall that the
order relation on~$\Pi_n$ is given as follows: if~$\pi,\sigma\in\Pi_n$,
then $\pi\leq\sigma$
if every block of~$\pi$ is contained in some block of~$\sigma$ (for short,
``$\pi$ refines~$\sigma$'').
In this problem,
you're going to calculate the number $\mu_n:=\mu_{\Pi_n}(\hat0,\hat1)$.

\probpart Calculate~$\mu_n$ by brute force for~$n=1,2,3,4$.
Make a conjecture about the value of~$\mu_n$ in general.    

\probpart Define a function~$f:\Pi_n\to\Qq[x]$ as follows: if~$X$
is a finite set of cardinality~$x$, then
  $$f(\pi) = \#\big\{h:[n]\to X \quad\big\vert\quad h(s)=h(s') \iff s,s'
    \text{ belong to the same block of }\pi\big\}.$$
For example, if~$\pi=\hat 1=\{\{1,2,\dots,n\}\}$ is the one-block
partition, then~$f(\pi)$ counts the constant functions from~$[n]$ to~$X$,
so~$f(\pi)=x$.  Find a formula for $f(\pi)$ in general.

\probpart Let~$g(\pi)=\sum_{\sigma\geq\pi} f(\sigma)$.  Prove that
$g(\pi) = x^{|\pi|}$ for all~$\pi\in\Pi_n$.  (Hint: What kinds of
functions are counted by the sum?)

\probpart Apply M\"obius inversion and
an appropriate substitution for~$x$ to calculate~$\mu_n$.

\end{document}
