\include{header}
\setboolean{answers}{true}
\newcounter{exprobno}
\setcounter{exprobno}{0}
\newcommand{\exprob}{\vskip10bp\addtocounter{probno}{1}{\bf(E{\arabic{probno}})}\quad}
\newcommand{\booksection}[1]{\bigskip\hrule\medskip\textbf{#1}}
\begin{document}

{\bf Math 724, Fall 2013\\
Homework \#5

Instructions:} Write up your solutions in LaTeX and hand in a hard copy in class on {\bf Friday, November~8.}  Collaboration is allowed (and in fact encouraged), but each student must write up his or her solutions independently and acknowledge all collaborators.

(\#1) Problem \#168.% moved from HW #4

(\#2) Problem \#170.% moved from HW #4

(\#3) Problem \#181.%(p.83)

(\#4) Problem \#195.%(p.88)

(\#5) Problem \#196.%(p.88)

(\#6) Problem \#197.%(p.89)

(\#7)  Problem \#205.

(\#8)  Problem \#206.

(\#9)  Problem \#207.

(Extra credit) Let $R=\Cc[x_1,\dots,x_n]$ be the ring\footnote{Quick definition of ``ring'': a set in which addition, subtraction and multiplication are well-defined operations that behave the way you think they ought to.} of polynomials in $n$ variables with complex coefficients.  As a vector space, $R$ decomposes as a direct sum
\[R = \bigoplus_{k=0}^\infty R_k = R_0\oplus R_1\oplus R_2\oplus\cdots\]
where $R_k$ means the vector subspace of polynomials that are homogeneous of degree $k$.  (This ``direct sum'' statement is just a fancy way of saying that every polynomial can be written in exactly one way as a constant plus a homogeneous linear form plus a homogeneous quadratic form plus yada yada yada.)  The \emph{Hilbert series} of $R$ is defined as the generating function
\[H_R(q)=\sum_{k\geq 0} (\dim R_k) q^k.\]

Finally, here's the problem: Find a nice closed-form expression for $H_R(q)$.

(Note: The Hilbert series can be defined for every graded ring, i.e., every ring~$R$ with a decomposition $R=\bigoplus_k R_k$ such that $R_kR_\ell\subseteq R_{k+\ell}$.  Graded rings are ubiquitous in algebraic combinatorics, and the Hilbert series of a graded ring is its most fundamental invariant.  For example, the quotient of $\Cc[x_1,\dots,x_n]$ by any homogeneous ideal (i.e., any ideal $I$ generated by homogeneous polynomials) is a graded ring; one remarkable fact is that the Hilbert series of $\Cc[x_1,\dots,x_n]/I$ is always a rational function of $q$.)

\end{document}