\include{header}
\begin{document}

{\bf Math 724, Fall 2013\\
Homework \#6

Instructions:} Write up your solutions in LaTeX and hand in a hard copy in class on {\bf Friday, November~22.}  Collaboration is allowed (and in fact encouraged), but each student must write up his or her solutions independently and acknowledge all collaborators.

\prob Consider the sequence $T_0,T_1,T_2,T_3,\dots$ defined recursively by
\[T_n = T_{n-1}+T_{n-2}+T_{n-3} \qquad\text{ for } n\geq 3.\]
Find a closed-form expression (in terms of $x,T_0,T_1,T_2$) for the generating function
\[\Omega = \sum_{n\geq0}T_nx^n.\]

\prob
Give a combinatorial interpretation for the coefficient of $q^k x^\ell$ in the power series
\[\prod_{n=1}^\infty (1+qx^n+qx^{2n}+qx^{3n}+qx^{4n}+\cdots).\]

\prob Bogart \#224.

\prob Bogart, Chapter 4, Supplementary Problem \#4.

\prob Bogart, Chapter 4, Supplementary Problem \#9.

\prob Bogart \#234.

\prob Bogart \#238 and \#239.  (Once you do \#238, problem \#239 should be easy.)

\prob The game of \textit{egdirb} uses a deck of 30 cards.  There are three suits: artichokes, ferrets, and pumpkins.  Each suit contains ten cards.  In one deal of egdirb, each of three players (Larry, Curly and Moe) receives a hand of 10 cards.  Use inclusion/exclusion to determine the probability that at least one player is dealt a void (i.e., zero cards) in at least one suit.
\bigskip

{\bf Extra credit:}\quad Redo problem \#7 for the game of bridge (with a standard four-suited deck and four players, each of whom receives 13 cards).  It is OK to write a computer program (preferably in Sage) to compute the answer; if you do so, include the source code in your written solutions.

\end{document}