\include{header}
\setboolean{answers}{false}
\begin{document}
\thispagestyle{empty}

\bf Math 724, Fall 2013\\
Take-Home Test \#2\rm

{\bf Instructions:} Write up your solutions using LaTeX.  
You may use books and notes, but you are not allowed to collaborate ---
you may not consult any human other than the instructor.
Solutions are due at the start of class on {\bf Friday, November 15.}
\medskip\hrule\bigskip

\prob
Let $n>0$ be an integer.

\Probpart{10}
How many labeled trees $T$ on vertex set $[n]$ have the property that the degree of every vertex is either 1 or 3?
Your answer should be a function of $n$ expressed without summation notation.

\Probpart{10}
Let $k$ be an integer with $0\leq k\leq n$.  How many labeled trees on vertex set $[n]$ have the property that vertices $1,2,3,\dots,k$ are all leaves
(i.e., each shares an edge with exactly one other vertex)?  (The tree can have other leaves as well.)  Your answer should be a function of $n$ and $k$ expressed without summation notation.

\vfill\Prob{20}
Let $S(k,n)$ denote Stirling numbers of the second kind.  Give a combinatorial proof that
\[S(k,n) = \sum_i \binom{k-1}{i-1} S(k-i,n-1)\]
for all positive integers $k,n$.  (By ``combinatorial,'' I mean ``explain why both sides of the equation count the same set of objects'' --- do not give a purely algebraic proof using, say, induction.)

\vfill\prob
Give combinatorial interpretations for the following numbers.  (In other words, describe what they count.)

\Probpart{10} The coefficient of $x^k$ in the infinite product
\[\prod_{n=1}^\infty (1+x^n+x^{2n}+\cdots+x^{n^2}).\]

\Probpart{10}
The coefficient of $q^\ell x^k$ in the infinite product
\[\prod_{n=1}^\infty \frac{1}{1-qx^n}.\]

\vfill\Prob{20}
Let $p,q$ be positive integers and let $C(p,q)$ denote the set of weak compositions of $p$ with $q$ parts.  Give an explicit bijection $C(p,q)\to C(q-1,p+1)$.

\vfill\Prob{20}
Recall that 1 Galleon is worth 17 Sickles and 1 Sickle is worth 29 Knuts.  Suppose that the Ministry introduces
a 3-Sickle and a 6-Knut piece (known respectively as a Trickle and a Hexknut).  With the new coinage,
how many ways are there of making change for a Galleon?  (If you are not an expert at Arithmancy, I recommend that you use Sage or another
computer algebra system to do the calculation.)

\end{document}
