\include{header}
\begin{document}

\textbf Math 724, Fall 2017\\
Homework \#1\\
Deadline: Friday, September 1, 5:00pm}

\textbf{Instructions:} Typeset your solutions in LaTeX.  Email your solutions to Jeremy (jlmartin@ku.edu) as a PDF file named with your last name and the problem set number (e.g., {\tt Abel1.pdf}).  Collaboration is encouraged, but each student must write up his or her solutions independently and acknowledge all collaborators.

\medskip\hrule\medskip

(\#1) Problem \#13.
\bigskip\bigskip

(\#2) Chapter 1 Supplementary Problem \#1.
\bigskip\bigskip

(\#3) Chapter 1 Supplementary Problem \#2.
\bigskip\bigskip

(\#4) Chapter 1 Supplementary Problem \#8.
Give at least two proofs, including at least one that is purely combinatorial (i.e., interpret both sides of the equation as different ways of counting the same set).
\bigskip\bigskip

(\#5)  A poker hand consists of five cards, drawn from a standard 52-card deck.

\quad (6a) How many different possible poker hands are there?

\quad (6b) A \emph{flush} is a hand with five cards of the same suit (e.g., $\spadesuit\mathsf{A8732}$ or $\clubsuit\mathsf{KJ984}$).  How many different flushes are there?

\quad (6c) A \emph{full house} is a hand with three cards of one rank and two of another (e.g., $\spadesuit\mathsf{A} \ \heartsuit\mathsf{8} \ \diamondsuit\mathsf{8} \ \clubsuit\mathsf{A8}$ or $\spadesuit\mathsf{63} \ \heartsuit\mathsf{63} \ \clubsuit\mathsf{6}$).  How many different full houses are there?

\bigskip\bigskip

(\#6)
Let $a_n=\sum_{k=0}^n\binom{n}{k}^2$. Calculate $a_n$ for $0\leq n\leq 3$.  Then stare at Pascal's triangle and make a conjecture about the value of $a_n$.  If you like, use Sage or another computer algebra system to check that your conjecture works for a few more values of $n$. Prove your conjecture \textit{combinatorially}; that is, find a set that can be counted in two ways so that one way of counting gives the formula for $a_n$, and the other way gives the formula you have conjectured.

\end{document}