\include{header}
\begin{document}
\thispagestyle{empty}

\textbf{Math 724, Fall 2017\\
Homework \#4\\
Deadline: Friday, October 13, 5:00pm}

\textbf{Instructions:} Typeset your solutions in LaTeX.  Email your solutions to Jeremy (jlmartin@ku.edu) as a PDF file named with your last name and the problem set number (e.g., {\tt Noether4.pdf}).  Collaboration is encouraged, but each student must write up his or her solutions independently and acknowledge all collaborators.

(\#1) Problem \#128.

(\#2) Problem \#129.

(\#3) Problem \#141.

(\#4) Problem \#142.
Implement your recurrence for (b) in Sage, and include the code you write in your submitted solutions.  (Use the {\tt verbatim} environment to get LaTeX to pay attention to line breaks and indentation.)  The \hreftext{http://jlmartin.faculty.ku.edu/math724/sage-vignette.pdf}{Sage Vignette} on the course website (also available as \hreftext{http://jlmartin.faculty.ku.edu/math724/sage-vignette.tex}{LaTeX source}) may be helpful; you may use or adapt the Sage code for Stirling numbers therein.  You can certainly use your code to answer part (c).

(\#5) Problem \#149.

(\#6) Problem \#150.  (Notice that this problem says ``onto'' where the previous one says ``to.''  So in \#149 you are counting all functions; in \#150 you are counting surjections.)

(\#7)  How many partitions $\lambda=(\lambda_1,\dots,\lambda_n)$ with $n$ parts have the property that $1\leq\lambda_k\leq n+1-k$ for all~$k$?  (Hint: Write down all such partitions for $n=1,2,3,4$, and you will have a pretty good guess of what the answer is.  Then find an appropriate bijection.

(Note: By definition, a partition is a weakly decreasing sequence of positive integers: $\lambda_1\geq\lambda_2\geq\cdots\geq\lambda_n>0$.  For instance, if $n=3$, the sequence $\lambda=(1,2,1)$ is not a partition even though it satisfies the criterion $1\leq\lambda_k\leq n+1-k$ for all $k$.)
\end{document}