\include{header}
\begin{document}
\thispagestyle{empty}

\textbf{Math 724, Fall 2017\\
Homework \#5\\
Deadline: Monday, October 30, 5:00pm}

\textbf{Instructions:} Typeset your solutions in LaTeX.  Email your solutions to Jeremy (jlmartin@ku.edu) as a PDF file named with your last name and the problem set number (e.g., {\tt Macaulay5.pdf}).  Collaboration is encouraged, but each student must write up his or her solutions independently and acknowledge all collaborators.

\vfill

\textbf{(\#1)} Bogart, Problem \#168.  (Parts (b), (c) and (d) don't require an explanation.  For the other parts, a couple of sentences should be enough.)

\vfill

\textbf{(\#2)} Bogart, Problem \#170.

\vfill

\textbf{(\#3)} Bogart, Problem \#197.

\vfill

\textbf{(\#4)} Bogart, Problem \#199, parts (a)--(f).  (The typesetting of this problem is horrible.  In the preamble, the last sentence should read, ``Then one possible picture of the tree $T$ is the product
\[P(T) = \prod_{\{i,j\}:\ \text{$i$ and $j$ are adjacent}} x_ix_j.''\]
Likewise, part (e) would be clearer with a displayed equation: ``\dots the picture enumerator
\[\sum{T:\ T\text{ is a tree on [n]}}P(T).''\]
Oh, and the second sentence in part (d) should end with a question mark.)

\vfill

\textbf{(\#5)} Bogart, Problems \#205--207 (they're really a package deal).

\vfill

\textbf{(\#6)} Let $R=\Cc[x_1,\dots,x_n]$ be the ring\footnote{Quick definition of ``ring'': a set in which addition, subtraction and multiplication are well-defined operations that behave the way you think they ought to.} of polynomials in $n$ variables with complex coefficients.  As a vector space, $R$ decomposes as a direct sum
\[R = \bigoplus_{k=0}^\infty R_k = R_0\oplus R_1\oplus R_2\oplus\cdots\]
where $R_k$ means the vector subspace of polynomials that are homogeneous of degree $k$.  (This ``direct sum'' statement is just a fancy way of saying that every polynomial can be written in exactly one way as a constant plus a homogeneous linear form plus a homogeneous quadratic form plus yada yada yada.)  The \emph{Hilbert series} of $R$ is defined as the generating function for the vector space dimensions of the graded pieces $R_k$:
\[H_R(q)=\sum_{k\geq 0} (\dim_\Cc R_k) q^k.\]

Find a nice closed-form expression for $H_R(q)$.  (Hint: You have already done this.  The problem really is about understanding the algebra well enough to recognize the appropriate combinatorics.)

(Note: The Hilbert series can be defined for every graded ring, i.e., every ring~$R$ with a decomposition $R=\bigoplus_k R_k$ such that $R_kR_\ell\subseteq R_{k+\ell}$.  Graded rings are ubiquitous in algebraic combinatorics, and the Hilbert series of a graded ring is its most fundamental invariant.  For example, the quotient of $\Cc[x_1,\dots,x_n]$ by any homogeneous ideal (i.e., any ideal $I$ generated by homogeneous polynomials) is a graded ring; one remarkable fact is that the Hilbert series of $\Cc[x_1,\dots,x_n]/I$ is always a rational function of $q$.)

\end{document}