\include{header}
\begin{document}
\thispagestyle{empty}

\textbf{Math 724, Fall 2017\\
Take-Home Test \#2\\
Deadline: Monday, November 13, 5:00pm}

\textbf{Instructions:} Typeset your solutions in LaTeX.  Email your solutions to Jeremy (jlmartin@ku.edu) as a PDF file named with your last name, e.g., {\tt HilbertTest2.pdf}.  You may refer to the textbook and your class notes, and you may cite the result of any problem from Chapter 1 assigned on Homeworks \#1--5 or done in class.  You may also use a computer algebra system such as Sage to carry out calculations and test conjectures.  However, \emph{you are not allowed to collaborate}; you may not consult any external resource or any human other than Jeremy.
\medskip\hrule

\prob
Let $n>0$ be an integer.

\Probpart{10}
How many labeled trees $T$ on vertex set $[n]$ have the property that the degree of every vertex is either 1 or 3?
Your answer should be a function of $n$ expressed without summation notation.

\Probpart{10}
Let $L\subseteq[n]$ and let $k=|L|$.  How many labeled trees $T$ on vertex set $[n]$ have the property that every element of $L$ is a leaf of $T$?  ($T$ can have other leaves as well.)  Your answer should be a function of $n$ and $k$ expressed without summation notation.

\vfill\Prob{20}
Let $S(k,n)$ denote Stirling numbers of the second kind.  Give a combinatorial proof that
\[S(k,n) = \sum_{i=1}^k \binom{k-1}{i-1} S(k-i,n-1)\]
for all positive integers $k,n$.  (By ``combinatorial,'' I mean ``explain why both sides of the equation count the same set of objects'' --- do not give a purely algebraic proof using, say, induction.)

\vfill\prob
Give combinatorial interpretations for the following numbers (i.e., describe what they count).

\Probpart{10} The coefficient of $x^k$ in the infinite product
\[\prod_{n=1}^\infty (1+x^n+x^{2n}+\cdots+x^{n^2}).\]

\Probpart{10}
The coefficient of $x^k$ in the infinite product
\[\prod_{n=1}^\infty \frac{1}{1-x^{n^2}}.\]

\Prob{20} Let $k,n$ be positive integers and let $P(k,n)$ denote the number of partitions of $k$ into $n$ parts (as in Bogart, p.~70).  Give a combinatorial proof that
\[P(k,1)+\cdots+P(k,n)=P(n+k,n).\]

\vfill\Prob{20}
Recall that 1 Galleon is worth 17 Sickles and 1 Sickle is worth 29 Knuts.  Suppose that the Ministry introduces
a 3-Sickle and a 6-Knut piece (known respectively as a Trickle and a Hexknut).  With the new coinage,
how many ways are there of making change for a Galleon?  (If you are not an expert at Arithmancy, I recommend that you use Sage or another
computer algebra system to do the calculation; if so, include the code you executed.)

\end{document}
