\include{header}
\begin{document}
\thispagestyle{empty}
{\bf Math 821, Spring 2018\\
Problem Set \#2\\
Deadline: Friday, February 16, 5:00pm}

\textbf{Instructions:} Typeset your solutions in LaTeX.  You are encouraged to use the \hreftext{http://jlmartin.faculty.ku.edu/math821/header.tex}{Math 821 header file}. Email your solutions to Jeremy (jlmartin@ku.edu) as a PDF file named with your last name and the problem set number (e.g., \texttt{Riemann2.pdf}).  Collaboration is encouraged, but each student must write up his or her solutions independently and acknowledge all collaborators.
\medskip
\hrule

\prob (Hatcher, p.20, \#22) Let $X$ be a finite graph lying in a half-plane $P\subset\Rr^3$ and intersecting the edge line $\ell$ of $P$ in a subset of its vertices.  Describe the homotopy type of the ``surface of revolution'' $Z$ obtained by rotating $X$ about $\ell$.  (Examples: If $X$ has two vertices and one edge, then $Z$ is a 2-sphere, a disk or a cylinder according as $X$ has two, one or zero vertices on $\ell$.  If $X$ has one vertex not on $\ell$ with a loop attached, then $Z$ is a torus.)


\prob Recall that for a space $X$ and basepoint $p\in X$, we have defined $\pi_1(X,p)$ to be
the set of homotopy classes of $p$-loops on $X$, or equivalently of continuous functions
$S^1\to X$.  Recall also that $S^0$ consists of two points (let's call them $a$ and $b$) with the discrete topology.
Accordingly, we could define $\pi_0(X,p)$ to be the set of homotopy classes of continuous functions $f:S^0\to X$
such that $f(a)=p$.  Describe the set $\pi_0(X,p)$ intrinsically in terms of $X$.  Is there a natural group structure on $\pi_0(X,p)$?
(Of course every set can be made into a group somehow, so the question is whether you can do it in a way depends only on, and tell you something about, the topology of $X$.)

\prob (Hatcher, p.38, \#2) Show that the change-of-basepoint homomorphism $\beta_h$ (see p.28)
depends only on the homotopy class of the path $h$.

\prob (Hatcher, p.38, \#7) Define $f:S^1\x I\to S^1\x I$ by $f(\theta,s)=(\theta+2\pi s,s)$,
so $f$ restricts to the identity on the two boundary circles of $S^1\x I$.   Show that $f$ is homotopic
to the identity by a homotopy $f_t$ that is stationary on \emph{one} of the boundary circles, but not
by any homotopy $f_t$ that is stationary on \emph{both} boundary circles.

\prob [Hatcher p.38 \#8] Does the Borsuk-Ulam theorem hold for the torus?  In other words, for every continuous $f:S^1\x S^1\to\Rr^2$ must there exist $(x,y)\in S^1\x S^1$ such that $f(x,y)=f(-x,-y)$?  Why or why not?

\prob [Hatcher p.39 \#12] Fix $p\in S^1$.  Show that every homomorphism $\pi_1(S^1,p)\to\pi(S^1,p)$ can be realized as the induced homomorphism $\phi_*$ for some continuous $\phi:S^1\to S^1$.

\prob [Hatcher, p.52, \#1]  Recall that the \defterm{center} of a group $G$ is defined as $Z(G)=\{g\in G:\ gh=hg \ \forall h\in G\}$.

\probpart Show that the free product $G*H$ of nontrivial groups $G$ and $H$ has trivial center.

\probpart Show that the only elements of $G*H$ of finite order are the conjugates of finite-order elements in~$G\cup H$.

\end{document}
