\include{header}
\begin{document}
\thispagestyle{empty}
{\bf Math 821, Spring 2018\\
Problem Set \#3 (Revised)\\
Deadline: Friday, March 9, 5:00pm}

\textbf{Instructions:} Typeset your solutions in LaTeX.  You are encouraged to use the \hreftext{http://jlmartin.faculty.ku.edu/math821/header.tex}{Math 821 header file}. Email your solutions to Jeremy (jlmartin@ku.edu) as a PDF file named with your last name and the problem set number (e.g., \texttt{Thurston3.pdf}).  Collaboration is encouraged, but each student must write up his or her solutions independently and acknowledge all collaborators.
\medskip
\hrule

\prob The \emph{dunce hat} is the space $D$ obtained from a triangle by identifying all three edges with each other, with the orientations indicated below.  Prove that $D$ is simply-connected (i) using Van Kampen's theorem; (ii) using what you know about 2-dimensional cell complexes.
\begin{center}
\begin{tikzpicture}[scale=2]
\draw[very thick,fill=pale] (0,0) -- (1,0) -- (.5,.866) -- cycle;
\draw[very thick, ->] (1,0) -- (.5,0);
\draw[very thick, ->] (1,0) -- (.75,.433);
\draw[very thick, ->] (.5,.866) -- (.25,.433);
\end{tikzpicture}
\end{center}

\prob (Hatcher, p.53, \#4, modified) Let $n\geq 1$ be an integer, and let $X\subset\Rr^3$ be the union of $n$ distinct rays emanating from the origin.  Compute $\pi_1(\Rr^3\sm X)$.

\prob Let $a_1,\dots,a_n$ be nonzero integers.  Construct a cell complex $X$ from $S^1$ as follows: For each $j=1,\dots,n$, attach a $2$-cell to $S^1$ by wrapping
it around the circle $a_j$ times.  Compute $\pi_1(X)$.

\prob (Hatcher, p.53, \#6, modified) Let $X$ be a path-connected cell complex, and let $Y$ be a cell complex obtained from $X$ by attaching an $n$-cell for some $n\geq 3$.  Show that the inclusion $X\inj Y$ induces an isomorphism $\pi_1(X)\isom\pi_1(Y)$.

\prob (Hatcher p.79, \#9) Show that if a path-connected, locally path-connected space $X$ has finite fundamental group, then every map $X\to S^1$ is nullhomotopic.  (Hint: Use the covering space map $\Rr\to S^1$.)

\prob (Hatcher p.80, \#12) Let $a$ and $b$ be the generators of $\pi_1(S^1\vee S^1,x_0)$ corresponding to the two copies of $S^1$, with $x_0$ their common point.  Draw a picture of the covering space $\tilde X$ of $S^1\vee S^1$ corresponding to the normal subgroup of $\pi_1(S^1\vee S^1)$ generated by $a^2$, $b^2$, and $(ab)^4$, and prove that this covering space is indeed the correct one.  (I.e., this group should be $p_*\pi_1(\tilde X,\tilde x_0)$.) 

\prob (Hatcher p.80, \#18) For a path-connected, locally path-connected, and semilocally simply-connected space~$X$, call a path-connected covering space $\tilde X\to X$ \emph{abelian} if it is normal and has abelian deck transformation group.  Show that $X$ has an abelian covering space $\hat X$ that is a covering space of every other abelian covering space of~$X$, and that such a ``universal abelian covering space'' is unique up to covering space isomorphism.  Describe this covering space explicitly for $X=S^1\vee S^1$ and $X=\bigvee^3 S^1$ (i.e., three circles, with a point from each identified).

\end{document}